```{r, echo = FALSE, include=FALSE}
set.seed(2018-12-08)
```

---

<center> <h1>A guide to using the DonkeyDosing package</h1> </center> 

---

## Introduction

---

The DonkeyDosing package is an R package that supports the anthelmintic dosing protocol being used by the Donkey Sanctuary, Devon.  There are two primary uses for this package:

1.  To help generate updated weather data in the correct format on a weekly basis throughout the grazing season.  This requires a source of weather data, which should be generated by scraping on a daily basis.

1.  To update the dosing tool every year in the spring.  This involves reading in historical data, new data from the previous year, weather data, and updating the coefficients used by the model.


---

## Generating weather data

---

There are a number of free services for scraping recent weather data for arbitrary coordinates. Until recently, DarkSky was preferred, but this service will close down by the end of 2022. One alternative is OpenWeather, although this is more limited than DarkSky in that data can only be obtained for the previous 5 days: this means that data must be scraped daily (or almost daily) in order to maintain a complete source of information.

A function is provided within the DonkeyDosing package to facilitate scraping.  The two input parameters are (1) the path to a folder where scraped data should be saved, and (2) an API key that has been set up via https://openweathermap.org/ (if this is omitted, then the key is assumed to be found in an "OpenWeatherKey.txt" file in the same folder as where the data is saved). The function can then be run as follows:

```{r}
ow_path <- "~/Documents/Resources/Datasets/DonkeyDosingWeather"

library("DonkeyDosing")
scrape_openweather(ow_path)
```


TODO

Planned content:

- Set up the model
- Import 2014-2017 locations, FEC and weather data from an archive
- Import 2018 locations and FEC data from the dosing tool
- Import 2018 weather data
- Run the model and obtain coefficients
- Graphical/check outputs
- Verification of Excel model outputs
- Clinical outputs (conditional modes of location/year)
- Scraping weather data
- Outputting animal/location and weather data ready for the 2019 dosing tool

---

## Vignette

This static vignette was built with:

```{r}
sessionInfo()
```

To rebuild the vignette, find the 'DonkeyDosing.Rmd' file in the 'vignette_source' directory (inside the inst directory of the package tarball), and (after editing as necessary) run:

```{r, eval=FALSE}
archivepath <- '..path_to_archive_data../archive.Rdata'
newpath <- '..path_to_2018_data../DosingTool2018.xlsx'

knitr::knit2html('DonkeyDosing.Rmd')
```

This will create an updated html file that can be used to replace the file provided in the vignettes folder of the package.

Note that the two necessary data files are not included in this R package, but were provided to the Donkey Sanctuary as part of the project ending in December 2018.
